%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DAXFS: A Delta-Log Memory-Native Filesystem with Speculative Branching
% USENIX Conference Paper
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix-2020-09}

\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,calc,patterns,decorations.pathreplacing}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xspace}
\usepackage{multirow}

\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  columns=flexible,
  language=C,
  morekeywords={__le32,__le64,u8,u16,u32,u64,cmpxchg},
}

\newcommand{\daxfs}{\textsc{DaxFS}\xspace}
\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
% Symbols drawn via TikZ to avoid loading cmsy10.pfb (corrupted in some environments)
\newcommand{\cmark}{\tikz[baseline=-0.5ex,scale=0.7]\draw[thick,green!60!black](0,0.1)--(0.1,0)--(0.3,0.25);}
\newcommand{\xmark}{\tikz[baseline=-0.4ex,scale=0.7]{\draw[thick,red!70!black](0,0)--(0.22,0.22);\draw[thick,red!70!black](0,0.22)--(0.22,0);}}

%-------------------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------------------

\date{}

\title{\Large \bf DAXFS: A Delta-Log Memory-Native Filesystem\\
  with Speculative Branching}

\author{
{\rm Cong Wang}\\
% Institution
} % end author

\maketitle

%-------------------------------------------------------------------------------
\begin{abstract}
%-------------------------------------------------------------------------------
We present \daxfs, a filesystem designed for byte-addressable
memory (persistent memory, CXL-attached memory, or DMA buffers) that
combines zero-copy direct access with a novel speculative branching
model.  \daxfs operates entirely through CPU load/store instructions on
DAX-mapped memory, eliminating the page cache, buffer heads, and data
copies that burden traditional filesystems.  Its core abstraction is a
read-only base image overlaid with \emph{delta-log branches}: each
branch records copy-on-write mutations as an append-only log that can
be committed to the root or discarded atomically.  Branches nest to
arbitrary depth, enabling N-level speculative execution where multiple
agents explore divergent paths and a single winner commits while all
siblings are invalidated.  Cross-kernel coordination uses hardware
\texttt{cmpxchg} on shared DAX memory, requiring no distributed
locking.  To our knowledge, \daxfs is the first filesystem to provide
hierarchical speculative branching with single-winner commit semantics
as a native primitive.  We describe the design, on-disk format,
branching semantics, and target use cases including AI agent
speculation, multikernel shared filesystems, CXL memory pooling, and
container rootfs sharing.
\end{abstract}

%-------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
%-------------------------------------------------------------------------------

Byte-addressable persistent memory~\cite{intel17:pmem,izraelevitz19:pmem}
and CXL-attached memory~\cite{cxl22:spec} have introduced a
fundamentally new tier in the memory hierarchy: storage-class media that
is directly accessible via CPU load and store instructions.  The Linux
DAX (Direct Access) subsystem~\cite{linuxdax} allows filesystems to
map this memory into user address spaces without copying data through
the page cache, offering the latency of DRAM with the capacity and
sharing properties of storage.

Despite this hardware capability, existing filesystems fail to exploit
it fully.  General-purpose DAX-aware filesystems like
NOVA~\cite{xu16:nova} and PMFS~\cite{dulloor14:pmfs} are designed for
persistent memory as a \emph{storage device} and retain significant
software overhead: metadata journaling, complex allocation, and
multi-level indexing structures.  Read-only filesystems like
EROFS~\cite{gao20:erofs} achieve simplicity but offer no mutation
path.  Union filesystems like OverlayFS~\cite{vangoor17:overlayfs}
provide copy-on-write layering but require page-cache-mediated I/O
and support only a single overlay level.  Log-structured filesystems
like NILFS2~\cite{konishi06:nilfs} and F2FS~\cite{lee11:f2fs} provide
snapshotting, but their checkpoints are strictly linear and cannot
branch into independent parallel paths.  COW-based filesystems like
Btrfs~\cite{rodeh13:btrfs} offer subvolume snapshots, but merging
independent subvolumes requires expensive tree diffing, and no
existing filesystem provides built-in ``discard all siblings on
commit'' semantics needed for speculative execution.

Meanwhile, emerging workloads demand a new combination of properties
that no existing filesystem offers.
AI agent systems perform \emph{speculative execution}: multiple agents
explore different solution paths on a shared environment, and only the
winning path is committed while losers are
discarded~\cite{wang26:branchfs,wei23:llmspec}.  Multikernel architectures run
independent OS instances on shared
hardware~\cite{baumann09:multikernel}, needing a common filesystem
without distributed locking.  Container deployments share a read-only
base image across hundreds of instances~\cite{merkel14:docker}, where
per-container copies waste memory.

We present \daxfs, a filesystem designed from the ground up for
byte-addressable memory.  Unlike prior persistent memory filesystems
that adapt block-device abstractions to byte-addressable media, \daxfs
treats memory as the primary storage medium and introduces three key
innovations:

\begin{enumerate}
\item \textbf{Memory-native zero-copy access.}  File reads resolve to
  direct memory loads with no page cache, no buffer heads, and no
  copies.  Unlike NOVA and PMFS, which still maintain per-inode logs
  and complex metadata trees even on byte-addressable memory, \daxfs
  uses a flat base image format that maps directly into the address
  space, reducing the read path to a single pointer dereference.

\item \textbf{Delta-log speculative branching.}  \daxfs introduces the
  first filesystem-level speculative branching primitive.  A read-only
  base image is overlaid with append-only delta logs, where each branch
  records mutations as log entries.  Branches nest to arbitrary depth,
  enabling N-level speculation with $O(\text{deltas})$ commit cost.
  This is fundamentally different from Btrfs snapshots (which require
  $O(\text{tree})$ diffing to merge), OverlayFS (which supports only
  one overlay level), and NILFS2 (which provides only linear
  checkpoints).

\item \textbf{Hardware-atomic cross-kernel coordination.}  Branch
  coordination across independent kernel instances uses
  \texttt{cmpxchg} directly on DAX memory, with no IPC, no distributed
  consensus protocol, and no kernel-to-kernel communication.  A single
  commit sequence counter in shared memory lets all participants
  detect state changes.  No prior filesystem achieves multi-kernel
  coordination without a distributed locking protocol or shared
  network service.
\end{enumerate}

Table~\ref{tab:feature-matrix} provides a feature-level comparison of
\daxfs against existing filesystem approaches across the dimensions
most relevant to our target workloads.

\begin{table*}[t]
\centering
\footnotesize
\begin{tabular}{@{}l@{\hspace{6pt}}c@{\hspace{6pt}}c@{\hspace{6pt}}c@{\hspace{6pt}}c@{\hspace{6pt}}c@{\hspace{6pt}}c@{\hspace{6pt}}c@{}}
\toprule
\textbf{Property} & \textbf{\daxfs} & \textbf{NOVA} & \textbf{PMFS} & \textbf{Btrfs} & \textbf{NILFS2} & \textbf{EROFS} & \textbf{OverlayFS} \\
\midrule
Zero-copy DAX reads        & \cmark & Partial & Partial & \xmark & \xmark & \xmark & \xmark \\
No page cache overhead     & \cmark & Partial & Partial & \xmark & \xmark & \xmark & \xmark \\
Hierarchical branching     & \cmark & \xmark & \xmark & Limited & \xmark & \xmark & 1-level \\
N-level speculation        & \cmark & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\
Atomic commit/abort        & \cmark & \xmark & \xmark & \xmark & \xmark & N/A & \xmark \\
Sibling invalidation       & \cmark & \xmark & \xmark & \xmark & \xmark & N/A & \xmark \\
Cross-kernel coordination  & \cmark & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\
No distributed locking     & \cmark & N/A & N/A & \xmark & N/A & N/A & N/A \\
Flat validatable format    & \cmark & \xmark & \xmark & \xmark & \xmark & \cmark & N/A \\
$O(\text{deltas})$ merge   & \cmark & N/A & N/A & \xmark & N/A & N/A & N/A \\
\bottomrule
\end{tabular}
\caption{Feature comparison of \daxfs with existing filesystems.
  ``Partial'' indicates DAX support exists but the filesystem still
  maintains significant metadata structures.  ``Limited'' indicates
  Btrfs supports snapshots but lacks commit/abort semantics and
  requires $O(\text{tree})$ merging.  ``1-level'' indicates OverlayFS
  supports a single upper layer over a read-only lower layer.}
\label{tab:feature-matrix}
\end{table*}

The rest of this paper is organized as follows.
Section~\ref{sec:motivation} motivates the design with target use
cases.  Section~\ref{sec:design} presents the \daxfs architecture and
its key innovations.
Section~\ref{sec:format} details the on-disk format.
Section~\ref{sec:branching} describes the branching model.
Section~\ref{sec:impl} covers implementation.
Section~\ref{sec:related} surveys related work in detail.
Section~\ref{sec:conclusion} concludes.

%-------------------------------------------------------------------------------
\section{Motivation}
\label{sec:motivation}
%-------------------------------------------------------------------------------

\daxfs targets scenarios where byte-addressable shared memory is
available and traditional filesystem abstractions impose unnecessary
overhead.  We identify four primary use cases that collectively
require a combination of properties unavailable in any existing
filesystem.

\paragraph{AI Agent Speculative Execution.}
Large language model agents increasingly employ speculative strategies:
multiple agents fork from a common state, explore different paths in
parallel, and the first successful result is committed while the
others are discarded.  This requires (1)~cheap forking of filesystem
state, (2)~isolated mutation within each fork, (3)~fast commit of the
winner, and (4)~atomic invalidation of losers.  No existing filesystem
offers all four properties as a unified primitive.  Btrfs
snapshots provide (1) and (2) but lack (3) and (4): merging a snapshot
back requires tree diffing, and there is no mechanism to atomically
invalidate sibling snapshots.  OverlayFS provides (1) and (2) for a
single level but cannot nest, and has no commit or invalidation
primitive.

\paragraph{Multikernel Shared Rootfs.}
In multikernel deployments~\cite{baumann09:multikernel}, independent
kernel instances share physical hardware.  A common root filesystem
eliminates per-kernel copies, but coordination is challenging: each
kernel has its own VFS cache, and distributed locking adds latency and
complexity.  CXL and persistent memory enable shared physical regions,
but existing filesystems require kernel-to-kernel communication for
consistency.  \daxfs solves this by placing the coordination state
directly in the shared DAX memory, using hardware atomics instead of
network protocols.

\paragraph{CXL Memory Pooling.}
CXL 3.0~\cite{cxl22:spec} enables memory pooling across hosts.  A
shared filesystem image in pooled CXL memory allows multiple hosts to
access common data without network transfers.  The filesystem must
support concurrent readers with independent write branches and must
coordinate using only hardware atomics on the shared memory region.
Traditional network filesystems (NFS, distributed Btrfs) add
unnecessary protocol layers when the memory is already directly
accessible.

\paragraph{Container Rootfs Sharing.}
Container deployments share a read-only base image (e.g., a Linux
distribution) across hundreds of containers~\cite{merkel14:docker,harter11:aufs}.
Current approaches
(tmpfs, OverlayFS) either duplicate the image per container or impose
page cache overhead.  \daxfs provides a single shared base image with
lightweight per-container branches that add only the delta of each
container's mutations, with no page cache copies and no per-container
memory duplication.

\subsection{Limitations of Existing Approaches}

Table~\ref{tab:comparison} summarizes why existing approaches fall
short for these use cases.  The fundamental gap is that no existing
filesystem combines direct memory access, hierarchical branching,
and cross-kernel coordination in a single design.

\begin{table}[t]
\centering
\small
\begin{tabular}{@{}lp{5.2cm}@{}}
\toprule
\textbf{Filesystem} & \textbf{Limitation} \\
\midrule
tmpfs/ramfs & Per-instance copies; $N$ containers = $N$ images in memory \\
OverlayFS & Single-level overlay only; copy-up on write; page cache overhead \\
EROFS & Read-only; no branching; fscache is per-kernel so sharing requires $N$ cache copies \\
famfs & Per-file allocation complexity; no self-contained images \\
cramfs & Block I/O plus page cache; no direct memory mapping \\
Btrfs & COW B-tree snapshots require $O(\text{tree})$ merge; no commit/abort semantics \\
NILFS2 & Linear checkpoints only; cannot branch independently \\
NOVA/PMFS & Complex metadata (B-trees, journals); no branching primitives \\
\bottomrule
\end{tabular}
\caption{Existing filesystem limitations for memory-native speculative
  branching workloads.  Each system lacks one or more of the core
  properties \daxfs provides: zero-copy DAX reads, hierarchical
  branching, and cross-kernel coordination.}
\label{tab:comparison}
\end{table}

%-------------------------------------------------------------------------------
\section{Design}
\label{sec:design}
%-------------------------------------------------------------------------------

\daxfs is designed around three principles: (1)~\emph{memory is the
storage}, not a cache for a block device; (2)~\emph{reads should be
free}, resolving to direct pointer dereferences; and (3)~\emph{writes
are speculative}, recorded in discardable delta logs rather than
applied in place.  These principles lead to a design that diverges
significantly from prior persistent memory filesystems.

\subsection{Architecture Overview}

Figure~\ref{fig:arch} illustrates the \daxfs architecture.  The
filesystem operates on a contiguous DAX-mapped memory region containing
a superblock, branch table, base image, optional page cache, and delta
region.

\begin{figure}[t]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[
    block/.style={draw, minimum width=3.2cm, minimum height=0.55cm,
      font=\small\sffamily, align=center},
    region/.style={draw, minimum width=3.2cm, minimum height=0.55cm,
      font=\small\sffamily, fill=blue!8},
    delta/.style={draw, minimum width=1.4cm, minimum height=0.55cm,
      font=\small\sffamily, fill=orange!15},
    arr/.style={-{Stealth[length=2mm]}, thick},
  ]
  % Memory layout
  \node[region] (sb) at (0, 0) {Superblock (4KB)};
  \node[region, below=1pt of sb] (bt) {Branch Table};
  \node[region, below=1pt of bt, fill=green!10] (base) {Base Image (RO)};
  \node[region, below=1pt of base, fill=yellow!10] (pc) {Page Cache (opt.)};
  \node[region, below=1pt of pc, fill=orange!10] (dr) {Delta Region};

  % Brace for DAX region
  \draw[decorate, decoration={brace, amplitude=5pt, mirror}]
    ([xshift=3pt]sb.north east) -- ([xshift=3pt]dr.south east)
    node[midway, right=8pt, font=\small\sffamily] {DAX Memory};

  % Branch details
  \node[delta, right=2.2cm of dr, yshift=0.4cm] (d1) {Branch 1};
  \node[delta, below=1pt of d1] (d2) {Branch 2};
  \node[delta, below=1pt of d2] (d3) {Branch N};
  \draw[arr, dashed] (dr.east) -- (d1.west);

  % Mount points
  \node[block, left=1.8cm of base, fill=green!8, yshift=0.3cm]
    (m1) {Mount /mnt};
  \node[block, left=1.8cm of dr, fill=orange!8, yshift=0.2cm]
    (m2) {Mount /spec};
  \draw[arr] (m1.east) -- (base.west);
  \draw[arr] (m2.east) -- (dr.west);
\end{tikzpicture}%
}
\caption{\daxfs memory layout.  A contiguous DAX region contains the
  superblock, branch table, read-only base image, optional shared page
  cache, and delta region.  Each mount binds to a specific branch.}
\label{fig:arch}
\end{figure}

A \daxfs instance starts with a \emph{base image}: a read-only
snapshot of a directory tree serialized in a flat format.  The base
image is immutable after creation.  All subsequent mutations occur in
\emph{branches}, each of which maintains an append-only delta log in
the delta region.

Each mount point is bound to exactly one branch.  The main branch
(parent~ID~0) is read-only and provides the base image.  To write, a
user creates a child branch, which inherits the parent's state and
records mutations in its own delta log.  This separation of immutable
base from mutable deltas is the foundation of the speculative
branching model, and it distinguishes \daxfs from filesystems like
NOVA, PMFS, or Btrfs, where all data is read-write from the start.

\subsection{Zero-Copy Read Path}
\label{sec:zero-copy}

In \daxfs, file reads resolve to direct pointers into DAX memory
without any intermediate copying.  The resolution path is:

\begin{enumerate}
\item Look up the inode in the branch's delta index.  If the inode has
  write extents in the current branch's delta log, return a pointer to
  the delta data.
\item Walk up the branch chain (child $\to$ parent $\to$ \ldots $\to$
  main), checking each ancestor's delta index.
\item If no delta covers the requested range, return a pointer into
  the base image's data area.
\end{enumerate}

At no point is data copied.  For \texttt{mmap}, the filesystem
installs PFN (page frame number) mappings directly to the DAX memory,
so user-space loads resolve to the physical memory in a single page
table walk.  By contrast, PMFS traverses a B-tree and NOVA walks a
per-inode log before returning data, even though both support DAX.

\subsection{Delta-Log Write Path}

Writes append entries to the branch's delta log using a sequential
bump allocator:

\begin{enumerate}
\item Allocate space in the delta log (a single atomic increment of
  the allocation pointer).
\item Write the delta header (type, inode, size, timestamp) and
  payload (file data, directory entry, etc.).
\item Update the in-memory index (red-black tree) to point to the new
  entry.
\end{enumerate}

Delta entries are typed: \texttt{WRITE} (file data), \texttt{CREATE}
(new file), \texttt{DELETE} (tombstone), \texttt{TRUNCATE},
\texttt{MKDIR}, \texttt{RENAME}, \texttt{SETATTR}, and
\texttt{SYMLINK}.  Each entry is self-describing and
variable-length.  The append-only structure ensures crash consistency
for individual operations without the journaling or copy-on-write
trees used by NOVA, PMFS, and Btrfs.

\subsection{Backing Store Mode}

For deployments where the base image is too large for DAX memory,
\daxfs supports a \emph{split mode}: metadata and delta logs reside in
DAX memory, while file data lives in an external backing file.  A
shared page cache in the DAX region provides cross-kernel caching of
backing-file pages using a three-state machine
(\textsc{Free}$\to$\textsc{Pending}$\to$\textsc{Valid}) with atomic
\texttt{cmpxchg} transitions.  Unlike the Linux kernel's page cache,
which is per-kernel and requires no cross-kernel sharing, this cache
resides in shared DAX memory and serves all kernel instances
simultaneously without any distributed cache coherence protocol.

%-------------------------------------------------------------------------------
\section{On-Disk Format}
\label{sec:format}
%-------------------------------------------------------------------------------

\daxfs uses a deliberately simple flat format (version~5) designed for
safe handling of untrusted images and efficient direct access.  This
simplicity is a conscious design choice: the format prioritizes
ease of validation and direct-mapped access over the space efficiency of
tree-based formats like Btrfs or the compression support of EROFS.

\subsection{Region Layout}

The memory region is divided into five areas, laid out sequentially:

\begin{description}
\item[Superblock (4\,KB).]  Magic number (\texttt{0x64617835},
  ``dax5''), version, flags, total size, and offsets to all other
  regions.  Contains a 64-byte global coordination structure for
  cross-kernel synchronization.

\item[Branch table.]  Up to 256 entries of 128 bytes each.  Each
  entry records the branch ID, parent ID, delta log offset and
  capacity, state (\textsc{Free}, \textsc{Active}, \textsc{Committed},
  \textsc{Aborted}), reference count, generation counter, and a
  32-character name.

\item[Base image.]  A self-contained read-only filesystem image with a
  4\,KB header, fixed-size inode table (64 bytes per inode), and data
  area containing file contents and flat directory entry arrays.

\item[Page cache (optional).]  Used in backing-store mode.
  Direct-mapped hash slots with 16-byte metadata each, managed by a
  clock-sweep eviction algorithm.

\item[Delta region.]  Contiguous space divided among active branches.
  Each branch owns a slice for its append-only delta log.
\end{description}

\subsection{Flat Directory Format}

Directories store fixed-size entries of 271 bytes each (the
\texttt{daxfs\_dirent} structure), containing a 2-byte name length, an
inode number, a mode field, and a 255-byte inline name.  This flat
design offers several security and performance advantages over the
pointer-based directory formats used by other filesystems (e.g., the
htree directories in ext4, the B-tree directories in Btrfs, or the
linked-list directories in tmpfs):

\begin{itemize}
\item \textbf{No pointer chasing.}  Directory entries are stored as
  flat arrays.  There are no linked lists, B-trees, or hash tables
  that could contain cycles or dangling pointers.

\item \textbf{Bounded iteration.}  The fixed entry size means
  directory traversal is a simple loop with a known upper bound,
  preventing denial-of-service via crafted deep structures.

\item \textbf{Inline names.}  File names are stored directly in the
  entry, eliminating string table indirection and the associated class
  of out-of-bounds vulnerabilities.

\item \textbf{Trivial validation.}  The entire format can be validated
  in a single pass at mount time (enabled by the \texttt{validate}
  mount option) without complex tree traversal.  By contrast,
  validating a Btrfs image requires traversing multiple B-trees, and
  EROFS validation must check compressed cluster boundaries.
\end{itemize}

\subsection{Delta Log Entries}

Each delta log entry begins with a 16-byte header:

\begin{lstlisting}
struct daxfs_delta_hdr {
  u8   type;       /* entry type */
  u8   reserved[3];
  __le32 total_size; /* header + payload */
  __le64 ino;        /* affected inode */
};
\end{lstlisting}

Eight entry types cover the full set of filesystem mutations:
\textsc{Write} carries an offset, length, and inline data (page-aligned
for mmap);
\textsc{Create} and \textsc{Mkdir} record a parent inode, new inode,
mode, and inline name;
\textsc{Delete} is a tombstone with parent inode and name;
\textsc{Truncate} records a new size;
\textsc{Rename} carries old and new parent inodes and names;
\textsc{Setattr} records changed attributes with a validity bitmask;
and \textsc{Symlink} stores both the link name and target.

This log format is deliberately more compact and simpler than the
per-inode logs used by NOVA (which maintain separate log pages with
forward/backward pointers) or the journal entries used by PMFS.
Because \daxfs delta entries are append-only and self-contained,
recovery requires no log replay or checkpoint restoration.

%-------------------------------------------------------------------------------
\section{Branching Model}
\label{sec:branching}
%-------------------------------------------------------------------------------

The branching model is the central innovation of \daxfs.  It provides
speculative execution semantics at the filesystem level: cheap fork,
isolated mutation, fast commit, and atomic abort.  No prior filesystem
provides this combination of properties.

\subsection{Branch Lifecycle}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    state/.style={draw, rounded corners, minimum width=1.6cm,
      minimum height=0.6cm, font=\small\sffamily},
    arr/.style={-{Stealth[length=2mm]}, thick},
  ]
  \node[state, fill=gray!15] (free) at (0,0) {Free};
  \node[state, fill=green!15] (active) at (3,0) {Active};
  \node[state, fill=blue!15] (committed) at (6,1) {Committed};
  \node[state, fill=red!15] (aborted) at (6,-1) {Aborted};

  \draw[arr] (free) -- node[above, font=\scriptsize] {create} (active);
  \draw[arr] (active) -- node[above, font=\scriptsize, sloped] {commit} (committed);
  \draw[arr] (active) -- node[below, font=\scriptsize, sloped] {abort / sibling} (aborted);
\end{tikzpicture}
\caption{Branch state transitions.  Branches move from \textsc{Free}
  to \textsc{Active} on creation, then to \textsc{Committed} (winner)
  or \textsc{Aborted} (loser or explicit abort).}
\label{fig:states}
\end{figure}

Figure~\ref{fig:states} shows the branch state machine.  A branch
begins in the \textsc{Free} state and transitions to \textsc{Active}
when created.  An active branch transitions to \textsc{Committed} when
its delta chain is merged to the root, or to \textsc{Aborted} when
explicitly aborted or when a sibling branch commits.

\subsection{N-Level Speculation}

A distinctive feature of \daxfs is that branches nest to arbitrary
depth, forming a tree rooted at the main branch.  Each branch inherits
its parent's state (the union of the parent's delta log and the base
image) and records its own mutations independently.  This contrasts
with existing approaches:

\begin{itemize}
\item \textbf{OverlayFS} supports only a single upper layer;
  nesting requires stacking mounts with compounding overhead.

\item \textbf{Btrfs} snapshots are independent trees, not a
  parent-child hierarchy; merging requires \texttt{btrfs send/receive}.

\item \textbf{NILFS2} checkpoints are strictly sequential, with no
  concept of independent parallel branches.
\end{itemize}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    branch/.style={draw, rounded corners, minimum width=1.1cm,
      minimum height=0.5cm, font=\small\sffamily},
    arr/.style={-{Stealth[length=2mm]}, thick},
    level distance=1.2cm,
    sibling distance=1.6cm,
  ]
  \node[branch, fill=green!10] (main) {main};
  \node[branch, fill=orange!10, below left=0.8cm and 0.6cm of main] (b1) {spec1};
  \node[branch, fill=orange!10, below right=0.8cm and 0.6cm of main] (b2) {spec2};
  \node[branch, fill=yellow!10, below left=0.8cm and 0.2cm of b1] (b1a) {spec1a};
  \node[branch, fill=yellow!10, below right=0.8cm and 0.2cm of b1] (b1b) {spec1b};

  \draw[arr] (main) -- (b1);
  \draw[arr] (main) -- (b2);
  \draw[arr] (b1) -- (b1a);
  \draw[arr] (b1) -- (b1b);
\end{tikzpicture}
\caption{N-level branch tree.  \texttt{spec1a} and \texttt{spec1b}
  are nested speculations within \texttt{spec1}.  Committing
  \texttt{spec1a} merges the chain spec1a$\to$spec1$\to$main and
  invalidates \texttt{spec1b} and \texttt{spec2}.}
\label{fig:tree}
\end{figure}

This model maps naturally to AI agent speculation.  An orchestrator
forks $N$ agents from a common state; each agent may itself fork
sub-agents for deeper exploration.  Figure~\ref{fig:tree} shows a
two-level speculation tree.

\subsection{Commit Semantics}

Committing a branch merges the \emph{entire chain} from the branch to
the root:

\begin{enumerate}
\item Acquire the global coordination lock (a \texttt{cmpxchg}-based
  spinlock in the superblock).
\item For each level from the committing branch up to main:
  \begin{enumerate}
  \item Invalidate all sibling branches: set state to
    \textsc{Aborted}, increment generation counter, unmap PFN
    mappings.
  \item Copy the child's delta log into the parent's log and rebuild
    the parent's in-memory index.
  \item Mark the child as \textsc{Committed}.
  \end{enumerate}
\item Increment the global commit sequence counter.
\item Release the coordination lock.
\end{enumerate}

The cost is $O(D)$ where $D$ is the total number of delta entries in
the chain, proportional to the actual mutations rather than the
filesystem size.  This contrasts with COW-subvolume approaches where
merging requires $O(\text{tree})$ diffing regardless of how few
changes were made (Section~\ref{sec:related}).

\subsection{Abort and Invalidation}

Aborting a branch discards its entire chain back to main.  Sibling
branches are \emph{not} affected; they continue operating normally.
This asymmetry is intentional: commit is a ``this path won'' signal
that must invalidate alternatives, while abort is a ``this path
failed'' signal that leaves other explorations intact.

Processes accessing invalidated branches receive clear error signals:
\texttt{mmap} regions trigger \texttt{SIGBUS} (PFN mappings are
unmapped on invalidation), and file operations return
\texttt{ESTALE}.  This clean error signaling is essential for the
speculative execution model: losing agents can detect invalidation
immediately and terminate gracefully rather than silently reading stale
data.

Single-level backtracking is also supported: unmounting a branch
discards only that branch while preserving the parent chain.

\subsection{Comparison with COW Subvolumes}

Table~\ref{tab:delta-vs-cow} compares delta-log branches with
traditional COW subvolumes in detail.

\begin{table}[t]
\centering
\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Aspect} & \textbf{COW (Btrfs)} & \textbf{Delta-Log (\daxfs)} \\
\midrule
Create & Snapshot tree & Allocate log \\
Commit & Diff + apply, $O(\text{tree})$ & Append, $O(\text{deltas})$ \\
Abort & Delete snapshot & Discard log \\
N-level & Complex merge & Chain to root \\
Invalidate & Not supported & Atomic \\
Cross-kernel & Not supported & \texttt{cmpxchg} on DAX \\
\bottomrule
\end{tabular}
\caption{Delta-log branches vs.\ COW subvolumes.  The delta-log
  model is purpose-built for speculative branching with single-winner
  semantics.}
\label{tab:delta-vs-cow}
\end{table}

\subsection{Comparison with Log-Structured Approaches}

Table~\ref{tab:log-compare} compares \daxfs with other log-structured
or snapshot-capable filesystems on the dimensions most relevant to
speculative branching.

\begin{table}[t]
\centering
\small
\begin{tabular}{@{}lcccc@{}}
\toprule
& \textbf{\daxfs} & \textbf{NILFS2} & \textbf{Btrfs} & \textbf{F2FS} \\
\midrule
Log-structured    & Yes & Yes & No & Yes \\
In-memory index   & Yes & Yes & No & Yes \\
Hierarchical branches & Yes & No & Partial & No \\
Commit/abort semantics & Yes & No & No & No \\
\bottomrule
\end{tabular}
\caption{Comparison with log-structured and snapshot-capable
  filesystems.  Only \daxfs provides both hierarchical branches and
  commit/abort semantics.}
\label{tab:log-compare}
\end{table}

%-------------------------------------------------------------------------------
\section{Implementation}
\label{sec:impl}
%-------------------------------------------------------------------------------

\daxfs is implemented as a Linux kernel module.  The implementation
comprises approximately 6,500 lines of C for the kernel module and
2,800 lines for user-space tools.

\subsection{In-Memory Indexing}

At mount time (or after a delta merge), \daxfs builds two red-black
tree indices from the delta log:

\begin{description}
\item[Inode index.]  Keyed by inode number.  Each entry caches the
  latest attributes (size, mode, uid, gid) and maintains a list of
  write extents in newest-first order for fast data lookup.

\item[Directory entry index.]  Keyed by a composite of parent inode
  number and name hash.  Each entry records whether the name was
  created or deleted (tombstone).
\end{description}

Data resolution walks the branch chain from child to parent.
Per-inode write extent lookup is $O(w)$ where $w$ is the number of
writes to that inode in the branch.  Directory entry lookup is
$O(\log n)$ per branch via the red-black tree, with $O(d)$ branches to
check where $d$ is the nesting depth.  This is simpler than the
multi-level B-tree traversals required by Btrfs or the extent-tree
lookups in NOVA, because the index structure mirrors the delta log
rather than encoding the full filesystem state.

\subsection{DAX Integration}

\daxfs implements the \texttt{iomap} interface for DAX file access.
For reads, \texttt{iomap\_begin} returns the physical address of the
data, whether it resides in a delta log or the base image, and the DAX
subsystem maps it directly into the user's page tables.  For writes,
\texttt{iomap\_begin} allocates a new delta entry and returns its
address; the DAX subsystem performs the write via direct store
instructions.

The filesystem supports three memory backing modes:
\begin{itemize}
\item \textbf{Physical address.}  A raw physical address range is
  mapped via \texttt{memremap}.
\item \textbf{DAX device.}  A \texttt{/dev/dax} or DMA heap device
  provides the memory.
\item \textbf{DMA-buf.}  A DMA buffer file descriptor is passed via
  the new mount API (\texttt{fsconfig} with \texttt{FSCONFIG\_SET\_FD}),
  enabling GPU/accelerator integration.
\end{itemize}

The DMA-buf backing mode enables zero-copy filesystem access from
GPUs and accelerators: by accepting a DMA-buf file descriptor at mount
time, \daxfs allows accelerator memory to serve as the filesystem's
backing store, enabling direct device-to-device data sharing without
CPU-mediated copies.

\subsection{Cross-Kernel Coordination}

The global coordination structure occupies 64 bytes within the
superblock at a fixed offset:

\begin{lstlisting}
struct daxfs_global_coord {
  __le64 commit_sequence;
  __le64 last_committed_id;
  __le32 coord_lock; /* 0=free, 1=held */
};
\end{lstlisting}

The spinlock uses \texttt{cmpxchg} directly on the DAX memory address.
Within a single node, DAX memory is cache-coherent across all CPUs, so
the atomic operates with normal SMP semantics.  For CXL-attached
shared memory, CXL~3.0 provides hardware coherence mechanisms
(back-invalidation and snoop filters) that extend this model across
hosts, though the coherence guarantees depend on the specific hardware
and fabric configuration.  Each mount caches the last-seen commit
sequence; if the sequence advances unexpectedly, the mount checks
whether its branch has been invalidated.

This approach is distinct from the distributed locking
used by cluster filesystems like GFS2 or OCFS2, which require a
separate lock manager service.  Because the lock state resides
directly in the shared memory region, \daxfs coordination works across
kernel instances within a node and, where CXL hardware coherence is
available, across CXL-connected hosts, with no additional
infrastructure.

\subsection{Shared Page Cache}

In backing-store mode, a direct-mapped hash cache resides in the DAX
region.  Each slot has 16 bytes of metadata:

\begin{lstlisting}
struct daxfs_pcache_slot {
  __le64 state_tag; /* bits[1:0]=state,
                       bits[63:2]=tag */
  __le64 ref_bit;   /* clock sweep */
};
\end{lstlisting}

The three-state machine
(\textsc{Free}$\to$\textsc{Pending}$\to$\textsc{Valid})
uses \texttt{cmpxchg} for lock-free transitions.  On a cache miss, a
kernel atomically transitions the slot from \textsc{Free} to
\textsc{Pending}, fills the data from the backing file, and
transitions to \textsc{Valid}.  Eviction uses a clock-sweep algorithm
that checks \texttt{ref\_bit} before reclaiming \textsc{Valid} slots.

This shared page cache addresses a problem unique to multi-kernel
environments: the standard Linux page cache is
per-kernel, so $N$ kernel instances reading the same backing file
would normally maintain $N$ separate cache copies.  The \daxfs shared
page cache in DAX memory ensures that a page filled by one kernel is
immediately accessible to all others, with no inter-kernel
communication.

\subsection{Security and Validation}

The flat format enables complete validation in a single sequential
scan.  When the \texttt{validate} mount option is specified, \daxfs
checks:
\begin{itemize}
\item Superblock magic and version.
\item All inode table entries: bounds on data offsets and sizes.
\item All directory entries: name lengths, inode references.
\item All delta log entries: type-specific size constraints, payload
  bounds.
\end{itemize}

The absence of pointer-based structures (no B-trees, no linked lists,
no hash chains) eliminates an entire class of attacks: cycle injection,
dangling pointer exploitation, and unbounded traversal.  This is a
significant advantage for mounting untrusted filesystem images, a
scenario common in container environments.  By contrast, validating a
Btrfs image requires traversing multiple interconnected B-trees, and
ext4 validation must check extent trees, directory htrees, and journal
integrity.

\subsection{User-Space Tools}

\daxfs provides three user-space tools:

\begin{description}
\item[\texttt{mkdaxfs}.]  Creates a \daxfs image from a source
  directory.  Handles hard-link deduplication, offset calculation, and
  flat directory serialization.  Can allocate memory from DMA heaps,
  physical addresses, or write to a file.

\item[\texttt{daxfs-branch}.]  Manages branch lifecycle: create,
  commit, abort, and list operations.

\item[\texttt{daxfs-inspect}.]  Dumps image metadata for debugging:
  superblock fields, branch table, inode table, and delta logs.
\end{description}

%-------------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}
%-------------------------------------------------------------------------------

We organize related work into five categories.

\subsection{Persistent Memory Filesystems}

PMFS~\cite{dulloor14:pmfs} was the first filesystem designed for
persistent memory, using a lightweight B-tree for file indexing and
copy-on-write for consistency.  NOVA~\cite{xu16:nova} introduced
per-inode log structuring for persistent memory, achieving high
concurrency by eliminating the global journal.
Strata~\cite{kwon17:strata} splits the I/O path between a user-space
log and a kernel-space digest layer, while
SplitFS~\cite{kadekodi19:splitfs} intercepts POSIX calls to perform
data operations in user space and metadata operations in the kernel.

All of these systems treat persistent memory as a general-purpose
storage device with full read-write semantics.  They maintain complex
metadata structures (B-trees in PMFS, per-inode logs with radix trees
in NOVA, multi-layer logs in Strata) to support the full range of
POSIX operations.  \daxfs takes a different approach: by
restricting the base image to read-only and recording all mutations as
simple append-only delta logs, \daxfs trades general-purpose storage
flexibility for a strictly simpler data path and the ability to
support speculative branching.  None of these filesystems provide any
branching or snapshot primitive.

\subsection{Log-Structured Filesystems}

NILFS2~\cite{konishi06:nilfs} provides continuous snapshotting via a
log-structured design where the entire filesystem is written
sequentially.  Its checkpoints capture the full filesystem state at
each point in time and can be mounted as read-only snapshots.
F2FS~\cite{lee11:f2fs} uses a log-structured approach optimized for
flash storage with a multi-head logging scheme.

\daxfs shares the log-structured principle with these systems but
applies it at a different granularity.  In NILFS2, the log captures
the entire filesystem state, and checkpoints are strictly linear (each
checkpoint is derived from its predecessor).  It is impossible to
create two independent checkpoints that branch from the same point.
In \daxfs, each branch maintains its own independent log, enabling a
tree of parallel branches that can be committed or aborted
independently.  F2FS provides no snapshot or branching mechanism at
all.

\subsection{COW and Snapshot Filesystems}

Btrfs~\cite{rodeh13:btrfs} is the most feature-rich Linux filesystem
for snapshots, providing COW subvolumes that share underlying extents.
Creating a Btrfs snapshot is fast (it copies only the root tree node),
but the key limitation for speculative workloads is merging.  To
merge snapshot $A$ back into the main subvolume, one must use
\texttt{btrfs send} to compute the delta between $A$ and its parent,
then \texttt{btrfs receive} to apply it.  This is an $O(\text{tree})$
operation that must traverse the entire B-tree to find changed extents.
Furthermore, Btrfs has no built-in ``commit this branch and invalidate
all siblings'' primitive; each snapshot is an independent entity that
must be managed explicitly.

By contrast, \daxfs commit is an $O(\text{deltas})$ operation that
copies only the actual mutations from child to parent.  Sibling
invalidation is atomic and built into the commit path.  The delta-log
model was designed specifically for the speculative execution pattern
of fork-explore-commit/abort, which Btrfs snapshots were not designed
to support.

\subsection{Read-Only and Container Filesystems}

EROFS~\cite{gao20:erofs} is a read-only filesystem widely used for
Android system partitions and container images.  It achieves high read
performance through a simple on-disk format and optional compression
but offers no mutation path.
OverlayFS~\cite{vangoor17:overlayfs} is the standard Linux union
filesystem for containers, layering a writable upper directory over a
read-only lower directory.

\daxfs combines the read-only base efficiency of EROFS with
multi-level writable branches, eliminating the page cache overhead
that OverlayFS imposes.  While OverlayFS supports only a single upper
layer and requires copy-up (copying an entire file to the upper layer
on the first write), \daxfs delta logs record only the changed bytes.
Moreover, OverlayFS nesting is not officially supported and introduces
compounding overhead, while \daxfs branches nest to arbitrary depth
with constant overhead per level.  Lazy container image distribution~\cite{harter11:aufs} also relies on
union mount stacking, with the same single-level limitation.

\subsection{Speculative Execution in Filesystems}

Speculator~\cite{nightingale06:specfs} applied speculative execution
to distributed filesystems, predicting server responses to hide
network latency and rolling back on mispredictions.  Its form of
speculation is qualitatively different from that of \daxfs: Speculator
optimistically assumes a remote operation will succeed and buffers
side effects until confirmation arrives, whereas \daxfs provides
isolated branch environments where multiple agents genuinely execute
different code paths, with the filesystem arbitrating the outcome
(commit one winner, abort all losers).

Wang and Zheng~\cite{wang26:branchfs} formalize this
fork-explore-commit/abort model as a general OS primitive called the
\emph{branch context}, providing copy-on-write state isolation with
independent filesystem views and process groups.  Their BranchFS
prototype~\cite{branchfs:repo} demonstrates the concept via FUSE,
achieving branch creation in under 350~$\mu$s and commit overhead
under 1~ms.  \daxfs~\cite{daxfs:repo} complements this work by
providing the same branching semantics at the kernel level on
DAX-mapped memory, eliminating FUSE overhead and enabling cross-kernel
coordination via hardware atomics on shared memory.

%-------------------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
%-------------------------------------------------------------------------------

We have presented \daxfs, a filesystem that exploits byte-addressable
memory to provide zero-copy access and speculative branching.  By
combining a flat read-only base image with append-only delta-log
branches and hardware-atomic cross-kernel coordination, \daxfs enables
a class of workloads (AI agent speculation, multikernel shared
filesystems, CXL memory pooling, and container rootfs sharing) that
are poorly served by existing filesystem designs.

The key innovations of \daxfs are: (1)~the delta-log branching model,
which provides $O(\text{deltas})$ commit cost, arbitrary nesting
depth, and clean single-winner semantics, where prior approaches
(Btrfs, NILFS2) require either $O(\text{tree})$ merging or offer only
linear checkpoints; (2)~the truly zero-copy read path that
eliminates all intermediate metadata structures between the
application and the data, going beyond even existing DAX-aware
filesystems like NOVA and PMFS; and (3)~cross-kernel coordination via
\texttt{cmpxchg} on shared DAX memory, which eliminates distributed
locking protocols for co-located kernels and, where CXL hardware
coherence is available, extends to multi-host sharing.

\daxfs demonstrates that memory-native filesystem design, treating
memory as the storage medium rather than a cache, enables
fundamentally new abstractions that are impossible in block-device
filesystems.  As byte-addressable memory becomes ubiquitous through
persistent memory and CXL, we expect speculative branching to become
an essential filesystem primitive for AI agent workloads and
shared-memory computing.

%-------------------------------------------------------------------------------
\section*{Availability}
%-------------------------------------------------------------------------------

\daxfs is available as open-source software~\cite{daxfs:repo}.  The
implementation includes the Linux kernel module, \texttt{mkdaxfs}
image creation tool, \texttt{daxfs-branch} management tool, and
\texttt{daxfs-inspect} debugging tool.  The companion BranchFS
FUSE-based prototype is also available~\cite{branchfs:repo}.

%-------------------------------------------------------------------------------
\begin{thebibliography}{21}

\bibitem{baumann09:multikernel}
A.~Baumann, P.~Barham, P.-E. Dagand, T.~Harris, R.~Isaacs, S.~Peter,
  T.~Roscoe, A.~Sch\"upbach, and A.~Singhania.
\newblock The multikernel: A new {OS} architecture for scalable multicore
  systems.
\newblock In {\em ACM Symposium on Operating Systems Principles (SOSP)}, pages
  29--44, 2009.

\bibitem{branchfs:repo}
{BranchFS}.
\newblock {BranchFS}: {FUSE}-based filesystem for speculative branching.
\newblock \url{https://github.com/multikernel/branchfs}.

\bibitem{cxl22:spec}
{CXL Consortium}.
\newblock {Compute Express Link (CXL)} specification, revision 3.0, 2022.

\bibitem{daxfs:repo}
{DAXFS}.
\newblock {DAXFS}: Delta-log memory-native filesystem with speculative
  branching.
\newblock \url{https://github.com/multikernel/daxfs}.

\bibitem{dulloor14:pmfs}
S.~R. Dulloor, S.~Kumar, A.~Keshavamurthy, P.~Lantz, D.~Reddy, R.~Sankaran,
  and J.~Jackson.
\newblock System software for persistent memory.
\newblock In {\em ACM European Conference on Computer Systems (EuroSys)}, pages
  15:1--15:15, 2014.

\bibitem{gao20:erofs}
X.~Gao, M.~Dong, S.~Miao, W.~Du, C.~Yu, and H.~Chen.
\newblock {EROFS}: A compression-friendly readonly file system for
  resource-scarce devices.
\newblock In {\em USENIX Annual Technical Conference (ATC)}, pages 149--162,
  2019.

\bibitem{harter11:aufs}
T.~Harter, B.~Salmon, R.~Liu, A.~C. Arpaci-Dusseau, and R.~H.
  Arpaci-Dusseau.
\newblock Slacker: Fast distribution with lazy {Docker} containers.
\newblock In {\em USENIX Conference on File and Storage Technologies (FAST)},
  pages 181--195, 2016.

\bibitem{intel17:pmem}
{Intel Corporation}.
\newblock {Intel Optane DC} persistent memory.
\newblock {\em Intel Technology Brief}, 2019.

\bibitem{izraelevitz19:pmem}
J.~Izraelevitz, J.~Yang, L.~Zhang, J.~Kim, X.~Liu, A.~Memaripour, Y.~J. Soh,
  Z.~Wang, Y.~Xu, S.~R. Dulloor, J.~Zhao, and S.~Swanson.
\newblock Basic performance measurements of the {Intel Optane DC} persistent
  memory module.
\newblock {\em arXiv:1903.05714}, 2019.

\bibitem{kadekodi19:splitfs}
R.~Kadekodi, S.~K. Lee, S.~Kashyap, T.~Kim, A.~Kolli, and V.~Chidambaram.
\newblock {SplitFS}: Reducing software overhead in file systems for persistent
  memory.
\newblock In {\em ACM Symposium on Operating Systems Principles (SOSP)}, pages
  494--508, 2019.

\bibitem{konishi06:nilfs}
R.~Konishi, Y.~Amagai, K.~Sato, H.~Hifumi, S.~Kihara, and S.~Moriai.
\newblock The {Linux} implementation of a log-structured file system.
\newblock {\em ACM SIGOPS Operating Systems Review}, 40(3):102--107, 2006.

\bibitem{kwon17:strata}
Y.~Kwon, H.~Fingler, T.~Hunt, S.~Peter, E.~Witchel, and T.~Anderson.
\newblock Strata: A cross media file system.
\newblock In {\em ACM Symposium on Operating Systems Principles (SOSP)}, pages
  460--477, 2017.

\bibitem{lee11:f2fs}
C.~Lee, D.~Sim, J.~Hwang, and S.~Cho.
\newblock {F2FS}: A new file system for flash storage.
\newblock In {\em USENIX Conference on File and Storage Technologies (FAST)},
  pages 273--286, 2015.

\bibitem{linuxdax}
{Linux Kernel Documentation}.
\newblock Direct access for files ({DAX}).
\newblock \url{https://www.kernel.org/doc/Documentation/filesystems/dax.txt}.

\bibitem{merkel14:docker}
D.~Merkel.
\newblock {Docker}: Lightweight {Linux} containers for consistent development
  and deployment.
\newblock {\em Linux Journal}, 239:2, 2014.

\bibitem{nightingale06:specfs}
E.~B. Nightingale, P.~M. Chen, and J.~Flinn.
\newblock Speculative execution in a distributed file system.
\newblock In {\em ACM Symposium on Operating Systems Principles (SOSP)}, pages
  191--205, 2005.

\bibitem{rodeh13:btrfs}
O.~Rodeh, J.~Bacik, and C.~Mason.
\newblock {BTRFS}: The {Linux} {B-Tree} filesystem.
\newblock {\em ACM Transactions on Storage (TOS)}, 9(3):1--32, 2013.

\bibitem{vangoor17:overlayfs}
B.~K. R. Vangoor, V.~Tarasov, and E.~Zadok.
\newblock To fuse or not to fuse: Performance of user-space file systems.
\newblock In {\em USENIX Conference on File and Storage Technologies (FAST)},
  pages 59--72, 2017.

\bibitem{wang26:branchfs}
C.~Wang and Y.~Zheng.
\newblock Fork, explore, commit: {OS} primitives for agentic exploration.
\newblock {\em arXiv:2602.08199}, 2026.

\bibitem{wei23:llmspec}
J.~Wei, X.~Wang, D.~Schuurmans, M.~Bosma, B.~Ichter, F.~Xia, E.~Chi, Q.~Le,
  and D.~Zhou.
\newblock Chain-of-thought prompting elicits reasoning in large language models.
\newblock In {\em Advances in Neural Information Processing Systems (NeurIPS)},
  2022.

\bibitem{xu16:nova}
J.~Xu and S.~Swanson.
\newblock {NOVA}: A log-structured file system for hybrid volatile/non-volatile
  main memories.
\newblock In {\em USENIX Conference on File and Storage Technologies (FAST)},
  pages 323--338, 2016.

\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
